---
output: html_document
editor_options:
  chunk_output_type: console
execute:
  eval: false
---
# Data Acquisition

```{r}
# Load necessary libraries
library(dplyr)
library(tidyverse)
library(mapme.biodiversity)
library(terra)
library(geodata)
library(sf)
library(progressr)
#library(rstac)
library(httr)
library(wdpar)
library(tmap)
library(pbapply)

local_output <- "data_out"
data_dest <- "PA_impacts_fires/data/source"
```


## Vector data

### National Boundaries

```{r}
mdg_0 <- gadm(country = "MDG", level = 0, path = "data")
mdg_1 <- gadm(country = "MDG", level = 1, path = "data")
mdg_2 <- gadm(country = "MDG", level = 2, path = "data")
# Store it locally
dir.create(paste0(local_output, "/admin_boundary"), recursive = TRUE, showWarnings = FALSE)
writeVector(mdg_0, paste0(local_output, "/admin_boundary/MDG_GADM_admin0.gpkg"), overwrite=TRUE)
writeVector(mdg_1, paste0(local_output, "/admin_boundary/MDG_GADM_admin1.gpkg"), overwrite=TRUE)
writeVector(mdg_2, paste0(local_output, "/admin_boundary/MDG_GADM_admin2.gpkg"), overwrite=TRUE)
# put to s3
map2(
  paste0(local_output, "/admin_boundary/MDG_GADM_admin", 0:2, ".gpkg"),
  paste0(data_dest, "/admin_boundary/MDG_GADM_admin", 0:2, ".gpkg"),
  put_to_s3
)
```


### Protected Areas

The download of WDPA data requires Chromium browser to be installed.
Followed the two subsequent cells to install it.
```{bash}
# 1. Gain root access and add the non-snap Chromium repository
sudo add-apt-repository ppa:xtradeb/apps -y

# 2. Update the package list
sudo apt-get update

# 3. Install the actual Chromium binary (and dependencies)
sudo apt-get install -y chromium

chmod +x /home/onyxia/work/PA_impacts_fires/renv/library/linux-ubuntu-noble/R-4.5/x86_64-pc-linux-gnu/processx/bin/supervisor
```

```{r}
# 1. Point R to the new binary location
Sys.setenv(CHROMOTE_CHROME = "/usr/bin/chromium")

# 2. Set the 'No-Sandbox' flag (Critical for Docker/Onyxia)
library(chromote)
set_chrome_args(c(default_chrome_args(), "--no-sandbox", "--disable-dev-shm-usage"))

# 3. Verify it works - this should return a path without erroring
chromote::find_chrome()
```

Load and clean WDPA data.

```{r}
# download protected area data for Madagascar
# (excluding areas represented as point localities)
mad_raw_wdpa <- wdpa_fetch(
  "Madagascar", wait = TRUE, download_dir = rappdirs::user_data_dir("wdpar")
)
# clean the data with high geometry precision
mad_clean_wdpa <- wdpa_clean(
  mad_raw_wdpa, geometry_precision = 10000,
  retain_status = NULL
)

#Load the boundary data
mad_boundary <- vect(paste0("/vsis3/projet-betsaka/", data_dest, "/admin_boundary/MDG_GADM_admin0.gpkg")) %>%
  sf::st_as_sf() %>%
  # repair any geometry issues, dissolve the border, reproject to same
  # coordinate system as the protected area data, and repair the geometry again
  st_set_precision(1000) %>%
  sf::st_make_valid() %>%
  st_set_precision(1000) %>%
  st_combine() %>%
  st_union() %>%
  st_set_precision(1000) %>%
  sf::st_make_valid() %>%
  st_transform(st_crs(mad_clean_wdpa)) %>%
  sf::st_make_valid()

# clip Madagscar's protected areas to the coastline
mad_clean_wdpa <- mad_clean_wdpa %>%
  dplyr::filter(REALM == "Terrestrial") %>%
  st_intersection(mad_boundary) %>%
  rbind(
    mad_clean_wdpa %>%
      dplyr::filter(REALM == "Marine") %>%
      st_difference(mad_boundary)
  ) %>%
  rbind(mad_clean_wdpa %>% dplyr::filter(!REALM %in% c("Terrestrial", "Marine")))

# recalculate the area of each protected area
mad_clean_wdpa <- mad_clean_wdpa %>%
  mutate(AREA_KM2 = as.numeric(st_area(.)) * 1e-6) %>%
  st_transform(4326) %>%
  sf::st_make_valid() %>%
  mutate(
    SHORT_NAME = case_match(NAME,
    "Betampona"                                              ~ "Betampona",
    "Tsimanampesotse"                                        ~ "Tsimanampe",
    "Befotaka Midongy"                                       ~ "Befotaka",
    "Isalo"                                                  ~ "Isalo",
    "Mantadia"                                               ~ "Mantadia",
    "Mananara Nord"                                          ~ "Mananara",
    "Ranomafana"                                             ~ "Ranomafana",
    "Tsaratanana"                                            ~ "Tsarata.",
    "Masoala"                                                ~ "Masoala",
    "Baie de Baly"                                           ~ "Baie Baly",
    "Bemaraha"                                               ~ "Bemaraha",
    "Marojejy"                                               ~ "Marojejy",
    "Andringitra"                                            ~ "Andringi.",
    "Namoroka"                                               ~ "Namoroka",
    "Mikea"                                                  ~ "Mikea",
    "Makira"                                                 ~ "Makira",
    "Ankarafantsika"                                         ~ "Ankaraf.",
    "Andohahela"                                             ~ "Andohahela",
    "Montagne d'Ambre"                                       ~ "Mont Ambre",
    "Analamazaotra"                                          ~ "Analamaz.",
    "Analamerana"                                            ~ "Analamer.",
    "Zombitse Vohibasia"                                     ~ "Zombitse",
    "Zahamena"                                               ~ "Zahamena",
    "Complexe des lacs Ambondro et Sirave (CLAS)"            ~ "CLAS Lakes",
    "Marolambo"                                              ~ "Marolambo",
    "Zones Humides Ankarafantsika  (CLSA)"                   ~ "CLSA Ank.",
    "Analavelona"                                            ~ "Analavelon",
    "Allées des Baobabs"                                     ~ "Baobabs",
    "Bemarivo"                                               ~ "Bemarivo",
    "Bora"                                                   ~ "Bora",
    "Kasijy"                                                 ~ "Kasijy",
    "Maningoza"                                              ~ "Maningoza",
    "Ambohijanahary"                                         ~ "Ambohijan.",
    "Ambatovaky"                                             ~ "Ambatovaky",
    "Mangerivola"                                            ~ "Mangeri.",
    "Kalambatrika"                                           ~ "Kalamba.",
    "Tampoketsa Analamaitso"                                 ~ "Tampoketsa",
    "Manombo"                                                ~ "Manombo",
    "Ivohibe"                                                ~ "Ivohibe",
    "Ambohitantely"                                          ~ "Ambohita.",
    "Beza Mahafaly"                                          ~ "Beza Maha.",
    "Le Lac Alaotra : Les Zones Humides et Bassins Versants" ~ "Lac Alaotra",
    "Anjanaharibe_sud"                                       ~ "Anjanaha.",
    "Ankarana"                                               ~ "Ankarana",
    "Manongarivo"                                            ~ "Manongari.",
    "Marotandrano"                                           ~ "Marotandra",
    "Nosy Mangabe"                                           ~ "Nosy Manga",
    "Andranomena"                                            ~ "Andranome.",
    "Cap Sainte Marie"                                       ~ "Cap Ste Ma",
    "Analalava"                                              ~ "Analalava",
    "Réserve spéciale Pointe à Larrée"                       ~ "Pte Larree",
    "Kinkony Lake"                                           ~ "Kinkony Lk",
    "Site Bioculturel d'Antrema"                             ~ "Antrema",
    "Complexe Tsimembo Manambolomaty"                        ~ "Tsimembo",
    "Parc Alarobia"                                          ~ "Alarobia",
    "Zones humides de Bedo"                                  ~ "Bedo",
    "Rivière Nosivolo et affluents"                          ~ "Nosivo Af",
    "Lac Mandrozo"                                           ~ "Mandrozo",
    "Complexe Anjozorobe Angavo"                             ~ "Anjozorobe",
    "Ankodida"                                               ~ "Ankodida",
    "Corridor forestier Bongolava"                           ~ "Bongolava",
    "Nord fotaky"                                            ~ "Fotaky",
    "Mandena"                                                ~ "Mandena",
    "Menabe Antimena"                                        ~ "Menabe Ant",
    "Ambohitr'Antsingy Montagne des Français"                ~ "Mont Franc",
    "Amoron'i Onilahy"                                       ~ "Amoro Onil",
    "Ambatoatsinanana"                                       ~ "Ambatoatsi",
    "Réserve de Tampolo"                                     ~ "Tampolo",
    "Rivière Nosivolo"                                       ~ "Nosivolo",
    "Complexe Lac Forêt Ambondrobe"                          ~ "Lac Ambond.",
    "Complexe Zones Humides Mangoky Ihotry"                  ~ "Mangoky",
    "Beanka"                                                 ~ "Beanka",
    "Sahafina"                                               ~ "Sahafina",
    "Corridor Forestier Ambositra Vondrozo"                  ~ "Ambositra",
    "Massif d'Itremo"                                        ~ "Itremo",
    "Ambatofotsy"                                            ~ "Ambatofo.",
    "Ampotaka Ankorabe"                                      ~ "Ampotaka",
    "Ampanganandehibe-Behasina"                              ~ "Ampanga.",
    "Ampasindava"                                            ~ "Ampasindav",
    "Galoko Kalobinono"                                      ~ "Galoko Kal",
    "Oronjia"                                                ~ "Oronjia",
    "Massif d'Ibity"                                         ~ "Ibity",
    "Forêt Naturel de Petriky"                               ~ "Petriky",
    "Andreba"                                                ~ "Andreba",
    "COMATSA Sud"                                            ~ "COMATSA S",
    "Bemanevika"                                             ~ "Bemanevika",
    "Loky Manambato"                                         ~ "Loky Manam",
    "Andrafiamena Andavakoera"                               ~ "Andrafiame",
    "Complexe des Zones Humides de Bemanevika"               ~ "ZH Bemane.",
    "Zones humides de l'Onilahy"                             ~ "Onilahy",
    "Anjajavy"                                               ~ "Anjajavy",
    "Ivohiboro"                                              ~ "Ivohiboro",
    "Corridor Ankeniheny Zahamena"                           ~ "Ankeniheny",
    "Forêt Naturelle de Tsitongambarika"                     ~ "Tsitongamb",
    "Maromizaha"                                             ~ "Maromizaha",
    "Mangabe-Ranomena-Sahasarotra"                           ~ "Mangabe",
    "Analabe-Betanatanana"                                   ~ "Analabe",
    "Agnalazaha"                                             ~ "Agnalazaha",
    "Makirovana Tsihomanaomby"                               ~ "Makirovana",
    "Vohidava Betsimalao"                                    ~ "Vohidava",
    "Ankarabolava"                                           ~ "Ankarabola",
    "Agnakatrika"                                            ~ "Agnakatri.",
    "Mahimborondro"                                          ~ "Mahimboron",
    "COMATSA Nord"                                           ~ "COMATSA N",
    "Manjakatompo Ankaratra"                                 ~ "Ankaratra",
    "Zones humides d'Ambondrobe"                             ~ "ZH Ambond.",
    "Parc national Tsimanampesotse"                          ~ "PN Tsimana",
    "Torotorofotsy"                                          ~ "Torotorofo",
    "Ambatotsirongorongo"                                    ~ "Ambatotsir",
    "Vohidefo"                                               ~ "Vohidefo",
    "Angavo"                                                 ~ "Angavo",
    "Behara-Tranomaro"                                       ~ "Behara",
    "Sud-Ouest Ifotaky"                                      ~ "Ifotaky",
    "Ranobe PK 32"                                           ~ "Ranobe",
    "Ambohidray"                                             ~ "Ambohidray",
    "Ambararata Londa"                                       ~ "Ambararata",
    "Vohimana"                                               ~ "Vohimana",
    "Madiro Mirafy"                                          ~ "Madiro Mir",
    "APC Miary"                                              ~ "APC Miary",
    "Makay"                                                  ~ "Makay",
    "Lac Sofia"                                              ~ "Lac Sofia",
    "Ankafobe"                                               ~ "Ankafobe",
    "APC Vezo"                                               ~ "APC Vezo",
    "Beampingaratsy"                                         ~ "Beampinga.",
    "Ankirihitra"                                            ~ "Ankirihi.",
    "APC Agnala"                                             ~ "APC Agnala",
    "APC Loharano"                                           ~ "APC Lohara",
    "APC Lovasoa"                                            ~ "APC Lovaso",
    "APC Mikea Mitambatra"                                   ~ "APC Mikea",
    "Tsinjoarivo Ambalaomby"                                 ~ "Tsinjoariv",
    "Complexe Bobaomby"                                      ~ "Bobaomby",
    "Complexe des  AP Ambohimirahavavy Marivorahona"         ~ "Ambohimira",
    "Andrefana Dry Forests"                                  ~ "Andrefana",
    "Forêts humides de l’Atsinanana"                         ~ "Atsinanana",
    "Kirindy Mite"                                           ~ "Kirindy M",
    "Sahamalaza Iles Radama"                                 ~ "Sahamalaza",
    "Lokobe"                                                 ~ "Lokobe",
    "Complexe Zones Humides Mahavavy Kinkony"                ~ "Mahavavy K",
    "Nosy Antsoha"                                           ~ "Nosy Antso",
    "Tsinjoriake"                                            ~ "Tsinjoriak",
    "Mangroves de Tsiribihina"                               ~ "Tsiribihi.",
    "Bombetoka Beloboka"                                     ~ "Bombetoka",
    "Ankingabe"                                              ~ "Ankingabe",
    "Mangroves de la Baie d'Ambaro"                          ~ "Ambaro",
    .default = NAME
  ))

# visual check with tmap
tmap_mode("view")

tm_shape(mad_clean_wdpa) +
  tm_borders() +
  tm_fill("IUCN_CAT", fill.legend = tm_legend(title = "Catégorie IUCN"))

tm_shape(mad_clean_wdpa) +
  tm_borders() +
  tm_fill("STATUS", fill.legend = tm_legend(title = "Status"))

#Export the cleaned data
sf::st_write(
  mad_clean_wdpa,
  dsn = "data/source/WDPA_data/mad_clean_wdpa.gpkg",
  delete_dsn = TRUE
)
# put to s3
map2(
  paste0("data/source/WDPA_data/mad_clean_wdpa.gpkg"),
  paste0(data_dest, "/WDPA_data/mad_clean_wdpa.gpkg"),
  put_to_s3
)
```

## Raster data


```{r}
# Use the mapme package to download available resources
mapme_dest <- paste0("/vsis3/projet-betsaka", data_dest)
mapme_options(outdir = mapme_dest)

# Convert to sf
mdg_sf <- st_as_sf(mdg)

# The latter could have been written more simply with:
with_progress({
  mdg_sf <- mdg_sf %>%
    get_resources(
      get_mcd64a1(years = 2000:2024), # MODIS MCD64A1
      get_gfw_treecover(version = "GFC-2024-v1.12"), # Global forest cover 2000
      get_nasa_srtm(), # NASADEM HGT v001
      get_accessibility_2000(), # JRC accessibility maps from 2000
      get_worldpop(years = 2000)
    )
})
# get_teow(path = NULL) for biomes should work but https://github.com/mapme-initiative/mapme.biodiversity/issues/450

```


## Global land cover datasets

```{r}
# list of URLs for GLAD database land cover data for Madagascar
# https://storage.googleapis.com/earthenginepartners-hansen/GLCLU2000-2020/v2/download.html
year <- c("2000", "2005", "2010", "2015", "2020")
tiles <- c("10S_040E", "10S_050E", "20S_040E", "20S_050E")
download_df <- expand.grid(year = year, tile = tiles)
# Create folder to save files
dir.create(paste0("./", local_output, "/GLAD_landcover"), recursive = TRUE)
# Download loop
lapply(
  X = 1:nrow(download_df),
  FUN = function(i) {
    url = paste0("https://storage.googleapis.com/earthenginepartners-hansen/GLCLU2000-2020/v2/", download_df$year[i], "/", download_df$tile[i], ".tif")
    download.file(
      url,
      paste0(local_output, "/GLAD_landcover/GLAD_Landcover_", download_df$year[i], "_", download_df$tile[i], ".tif"),
      mode = "wb"
)})
# put on s3
map2(
  paste0(local_output, "/GLAD_landcover/GLAD_Landcover_", download_df$year, "_", download_df$tile, ".tif"),
  paste0(data_dest, "/GLAD_landcover/GLAD_Landcover_", download_df$year, "_", download_df$tile, ".tif"),
  put_to_s3
)

```

## World settlements footprint evolution data

```{r}
# List of tile coordinates (just the number pairs)
# list of tiles come from https://download.geoservice.dlr.de/WSF_EVO
tiles <- c(
  "46_-22", "44_-22", "42_-22", "42_-24", "42_-26", "44_-26",
  "46_-26", "44_-24", "46_-24", "48_-24", "48_-22", "48_-20",
  "44_-20", "42_-18", "44_-18", "46_-18", "46_-20", "48_-18",
  "50_-18", "50_-16", "48_-16", "46_-16", "44_-16", "44_-14",
  "46_-14", "48_-14", "50_-14", "48_-12"
)
urls <- paste0("https://download.geoservice.dlr.de/WSF_EVO/files/WSFevolution_v1_", tiles, ".tif")

# Create folder to save files
dir.create(paste0("./", local_output, "/WSF_EVO"), recursive = TRUE)

# Download loop
lapply(
  X = urls,
  function(url) {
  download.file(
    url,
    paste0(local_output, "/WSF_EVO/", str_extract(url,"WSFevolution_v1_(.*).tif")),
    mode = "wb"
  )
})
# put on s3
map2(
  paste0(local_output, "/WSF_EVO/WSFevolution_v1_", tiles, ".tif"),
  paste0(data_dest, "/WSF_EVO/WSFevolution_v1_", tiles, ".tif"),
  put_to_s3
)
```


### Fire data

Batch download data from MODIS, using Appears.
Necessary to be login into Earthdata.
To obtain the download list, a request must be done in Appears. Then, select all the files in the request, click download and save download list.
````{r}
# Batch download from Appears, require authentification

# Put on s3
modis_files <- list.files(paste0(local_output, "/MODIS_Appears"), recursive = TRUE)
map2(
  paste0(local_output, "/MODIS_Appears/", modis_files),
  paste0(data_dest, "/MODIS_Appears/", modis_files),
  put_to_s3
)
````

Fire CCI data from ESA.
````{r}
library(httr)
dir.create(paste0("./", local_output, "/MODIS_CCI"), recursive = TRUE)

for (year in 2019:2022) {
  for (month in str_pad(1:12, width = 2, pad = "0")) {
    if (year == 2007 | (year == 2019 & month >= 10)) {
      target_url = paste0("https://dap.ceda.ac.uk/neodc/esacci/fire/data/burned_area/MODIS/pixel/v5.1/compressed/", year, "/new-corrected/", year, month, "01-ESACCI-L3S_FIRE-BA-MODIS-AREA_5-fv5.1.tar.gz?download=1")
    }  else {
      target_url = paste0("https://dap.ceda.ac.uk/neodc/esacci/fire/data/burned_area/MODIS/pixel/v5.1/compressed/", year, "/", year, month, "01-ESACCI-L3S_FIRE-BA-MODIS-AREA_5-fv5.1.tar.gz?download=1")
    }
    filename <- sub("?download=1", "", basename(target_url), fixed = TRUE)
    save_path <- paste0("./", local_output, "/MODIS_CCI/", filename)
    response <- download.file(target_url, save_path)
    archive::archive_extract(save_path, paste0("./", local_output, "/MODIS_CCI"))
    unlink(save_path)
  }
}

# Put on s3
burn_files <- list.files(paste0(local_output, "/MODIS_CCI"), recursive = TRUE)
map2(
  paste0(local_output, "/MODIS_CCI/", burn_files),
  paste0(data_dest, "/MODIS_CCI/", burn_files),
  put_to_s3
)
````
